---
title: "Multi-core processing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F, error = F, fig.width = 9)
library(flextools)
library(dplyr)
library(lubridate)
library(ggplot2)
```

The calculations done by functions `smart_charging` or `optimize_demand` can be done by multiple time windows in parallel. At the same time, in computer science, a common practice to reduce computation times is parallel processing when the computer has multiple CPUs (i.e. multi-core processing). In parallel processing, a task is divided into multiple sub-tasks and assigned to multiple cores to make de calculations in parallel. For parallel processing in R we make use of the R base package `parallel`.

We can discover the number of physical CPUs in our computer with function `parallel::detectCores`:
```{r}
n_cores <- parallel::detectCores(logical = FALSE)
print(n_cores)
```

We set the parameter `logical = FALSE` because it is important to know the **physical** CPUs, not the logical (or virtual/thread) ones seen by the OS. More information about this parameter can be found in `parallel::detectCores` documentation.

Below you can find some demonstrations of parallel processing using flextools. Note that in all demonstrations, the optimal number of cores to use is around 6, since from this value no big improvement is achieved. So, remember, **more doesn't always mean better**.


### Smart charging

Below, an example of calculating `smart_charging` of 10.000 EV charging sessions with different number of cores:

```{r, echo = T, eval = F}
sessions <- evsim::california_ev_sessions_profiles %>%
  slice_head(n = 10000) %>%
  evsim::adapt_charging_features(time_resolution = 15)

sessions_demand <- evsim::get_demand(sessions, resolution = 15, mc.cores = n_cores)

opt_data <- tibble(
  datetime = sessions_demand$datetime,
  production = 0
)

cores_time <- tibble(
  cores = 1:n_cores,
  time = 0
)

for (mcc in cores_time$cores) {
  results <- system.time(
    smart_charging(
      sessions, opt_data, opt_objective = "grid", method = "curtail",
      window_days = 1, window_start_hour = 5, mc.cores = mcc
    )
  )
  cores_time$time[mcc] <- as.numeric(results[3])
}
```

```{r, echo=F, eval = T}
# saveRDS(cores_time, "multi-core-data/cores_time_sc.RDS")
cores_time <- readRDS("multi-core-data/cores_time_sc.RDS")
```


And below there's the plot of the results, were we can see that, for this computer, the optimal number of cores to use is 6, since from this value no big improvement is achieved:

```{r}
#| fig.alt: >
#|   Computation time for smart charging using multiple cores
cores_time %>%
  ggplot(aes(x = cores, y = time)) +
  geom_col() + 
  labs(x = "Number of cores", y = "Computation time (seconds)")
```



### Demand optimization

Below, an example of `optimize_demand` for the `building` profile with different number of cores:

```{r, echo = T, eval = F}
opt_data <- flextools::energy_profiles %>%
  rename(
    production = "solar",
    flexible = "building"
  )

cores_time <- tibble(
  cores = 1:n_cores,
  time = 0
)

for (mcc in cores_time$cores) {
  results <- system.time(
    opt_data %>%
      optimize_demand(
        opt_objective = 0.5,
        direction = "backward",
        flex_window_hours = 6,
        time_horizon = 12,
        mc.cores = mcc
      )
  )
  cores_time$time[mcc] <- as.numeric(results[3])
}
```

```{r, echo=F, eval = T}
# saveRDS(cores_time, "multi-core-data/cores_time_od.RDS")
cores_time <- readRDS("multi-core-data/cores_time_od.RDS")
```


And below there's the plot of the results, were we can see that, for this computer, the optimal number of cores to use is 6, since from this value no big improvement is achieved:

```{r}
#| fig.alt: >
#|   Computation time for demand optimization using multiple cores
cores_time %>%
  ggplot(aes(x = cores, y = time)) +
  geom_col() + 
  labs(x = "Number of cores", y = "Computation time (seconds)")
```




### Battery optimization

Below, an example of `add_battery_optimization` with different number of cores:

```{r, echo = T, eval = F}
opt_data <- flextools::energy_profiles %>%
  rename(
    production = "solar"
  ) %>%
  select(any_of(c(
    "datetime", "production", "building", "price_imported", "price_exported"
  ))) %>%
  mutate(
    static = building
  )

opt_battery <- 

cores_time <- tibble(
  cores = 1:n_cores,
  time = 0
)

for (mcc in cores_time$cores) {
  results <- system.time(
    opt_data %>%
      add_battery_optimization(
        opt_objective = 0.5,
        Bcap = 50, Bc = 4, Bd = 4,
        window_start_hour = 5,
        mc.cores = mcc
      )
  )
  cores_time$time[mcc] <- as.numeric(results[3])
}
```

```{r, echo=F, eval = T}
# saveRDS(cores_time, "multi-core-data/cores_time_ob.RDS")
cores_time <- readRDS("multi-core-data/cores_time_ob.RDS")
```


And below there's the plot of the results, were we can see that, for this computer, the optimal number of cores to use is 6, since from this value no big improvement is achieved:

```{r}
#| fig.alt: >
#|   Computation time for battery optimization using multiple cores
cores_time %>%
  ggplot(aes(x = cores, y = time)) +
  geom_col() + 
  labs(x = "Number of cores", y = "Computation time (seconds)")
```



