---
title: "Smart charging"
output: html_document
---


```{r setup, echo=FALSE, message=F, warning=FALSE, fig.width=12}
library(flextools)
# devtools::load_all()
library(dplyr)
library(lubridate)
library(dygraphs)
library(dutils)
library(ggplot2)
library(evsim)
library(purrr)
```
The smart charging concept is widely used in different fields and applications. In `flextools` package, we define *smart charging* as a tool to coordinate individual EV charging sessions in order to obtain the optimal aggregated demand profile according to a certain objective. There are different practical **methods** to coordinate each session, such as:

* **Postpone**: shifting charging start time over time
* **Interrupt**: stop charging during certain time
* **Curtail**: limiting charging power during certain time
* **Vehicle-to-grid (V2G)**: combine charge and discharge during connection time. Not available in flextools yet.

At the same time, the charging sessions can be coordinated with different objectives or goals, such as minimizing the interaction with the power grid, minimizing the energy cost, participating in flexibility or imbalance markets, not surpassing grid constraints or capacity limits, accomplishing with demand-response programs, etc. 

The function `smart_charging()` aims to provide a framework to simulate any of these situations to analyze the impact and benefits of electric vehicle flexibility. 


# Smart charging algorithm

We have divided the smart charging algorithm contained inside `smart_charging()` function in two different stages:

1. Setpoint calculation for the aggregated EV demand curve
2. Scheduling of the individual sessions to match the setpoint of the aggregated demand

Below, we will briefly explain the process performed in each one of theses steps.


## Setpoint calculation

The *setpoint* for the aggregated EV demand is understood as the desired, optimal or allowed power demand, depending on the objective and characteristics of the control over the charging points. 

Currently, `flextools` allows the following methods to define the setpoint of the aggregated EV demand:

* **Optimization setpoint**: a Quadratic programming optimization is performed to minimize the interaction with the power grid (see [Net power optimization](https://mcanigueral.github.io/flextools/articles/minimize_net_power.html)), the energy cost (see [Energy cost optimization](https://mcanigueral.github.io/flextools/articles/minimize_cost.html)) or both (see [Combined optimization](https://mcanigueral.github.io/flextools/articles/combined_optimization.html)), setting the parameter `opt_objective` accordingly. The `smart_charging()` function is making use of the `optimize_demand()` function internally, using parameters `direction="forward"` to postpone EV sessions and `time_horizon=NULL` to exploit their flexibility until the end of the optimization window.  
* **Grid capacity profile**: a maximum grid capacity profile can be used as a constrain to coordinate charging sessions below the limits. For this functionality, the grid capacity must be a variable (`grid_capacity`) in the `opt_data` parameter of the `smart_charging()` function, together with `opt_objective = "none"`.


## Scheduling algorithm

The smart charging method is defined by the parameter `method` in the `smart_charging()` function, which can be `"postpone"`, `"curtail"`, `"interrupt"` or `"none"`. If `method = "none"`, the sessions schedule is not modified and the calculated setpoints are returned as a optimal results. If `method` is different than `"none"`, after obtaining the setpoint $S_t$, the the scheduling algorithm follows the sequence below for every time slot $t$:

1. Calculate $V_t$, the EV demand
2. Get the time slots where $V_t > S_t$
3. Stop if no more flexibility is required
4. For every time slot where $V_t > S_t$

    4.1. Get the power difference between load and setpoint in the time slots where $V_t > S_t$
    
    4.2. Select flexible sessions (see definition below) and set a priority order
    
    4.3. Go to next time slot if no more flexibility is available
    
    4.4. Coordinate flexible sessions to match the setpoint
    
5. Stop if no more flexibility is available
6. Return the new schedule of EV sessions

To classify a connected EV as a **flexible session** or not, the `flextools` package defines the following conditions according to smart charging method used:

* Postpone: the EV has not started charging yet, and the energy required can be charged during the rest of the connection time at the nominal charging power. From all flexible sessions in a time slot, the ones connecting earlier will have priority over the later sessions.
* Interrupt: the charge is not completed yet, and the energy required can be charged during the rest of the connection time at the nominal charging power. From all flexible sessions in a time slot, the ones that have been charging during less time will have priority over the sessions that have been already charging (rotation system).
* Curtail: the charge is not completed yet, and the energy required can be charged during the rest of the connection time at a lower power than the nominal charging power.

The energy and power constraints of these conditions can also be defined in function `smart_charging()` with the parameters `energy_min` and `charging_power_min`, representing the minimum allowed ratios of the nominal values. With `energy_min = 0` the algorithm considers that EVs may disconnect without charging all their energy requirements, while with `energy_min = 1` the algorithm will make sure that all EV users charge their energy requirements, even though the setpoint is not achieved. On the other hand, `charging_power_min = 0` means that, when `method = "curtail"`, the charging power can be reduced until 0, so the session can stop, while a value of `charging_power_min = 0.5` would only allow curtailing the EV charging power until the half of its nominal charging power.

# Smart charging examples 

Below, some examples of `smart_charging()` are illustrated for both the **grid congestion simulation**, where we have a maximum grid capacity to not surpass, and the **optimization simulation**, where the net power usage or/and the energy cost are optimized thanks to smart charging.

First of all, we can simulate EV sessions using the `evsim` package, which provides the function `evsim::get_custom_ev_model` to create a custom EV model to later simulate EV sessions with `evsim::simulate_sessions()` function.

We can create an EV model with custom time-cycles and user profiles. In this case, we will consider just one EV user profile (see the [EV user profile concept](https://mcanigueral.github.io/evprof/articles/evprof.html) from package `evprof`), "HomeEV", which will represent a "charge-at-home" pattern starting in average at 18:00 until next morning.

```{r}
# For workdays time cycle
workdays_parameters <- dplyr::tibble(
  profile =  "HomeEV",
  ratio = 100,
  start_mean = 18,
  start_sd = 3,
  duration_mean = 13,
  duration_sd = 1,
  energy_mean = 15,
  energy_sd = 4
)

# For weekends time cycle
weekends_parameters <- dplyr::tibble(
  profile = "HomeEV",
  ratio = 100,
  start_mean = 12,
  start_sd = 4,
  duration_mean = 3,
  duration_sd = 2,
  energy_mean = 4,
  energy_sd = 4
)

parameters_lst <- list(workdays_parameters, weekends_parameters)

# Get the whole model
ev_model <- get_custom_ev_model(
  names = c("Workdays", "Weekends"),
  months_lst = list(1:12, 1:12),
  wdays_lst = list(1:5, 6:7),
  parameters_lst = parameters_lst,
  connection_log = FALSE,
  energy_log = FALSE,
  data_tz = "Europe/Amsterdam"
)
```

Once we have our own model, we can simulate EV sessions for a specific date (e.g. today):

```{r}
set.seed(1234)
ev_sessions <- simulate_sessions(
  evmodel = ev_model, 
  sessions_day = tibble(time_cycle = c("Workdays", "Weekends"), n_sessions = c(5, 5)),
  user_profiles = NULL,
  charging_powers = tibble(power = 3.7, ratio = 1), 
  dates = today(), 
  resolution = 15
)
```

```{r, echo=FALSE}
ev_sessions %>% knitr::kable()
```


Finally we can calculate the time-series power demand from each EV with function `evsim::get_demand()`, using parameter `by="Sessions"`:

```{r}
ev_demand <- ev_sessions %>% 
  get_demand(by = "Session") 
```
```{r, echo=FALSE}
ev_demand %>% 
  dyplot(ylab = "Power demand (kW)", stackedGraph = T, fillGraph = T)
```


It is visible that these EV users tend to coincide during the evening peak hours, while coordinated they could make a more efficient usage of the power grid. Following, you will find some examples about how to coordinate these EV sessions using `smart_charging()`.

To check the results, the object returned by function `smart_charging()` is always list with 3 more objects:

* **setpoints**: time-series data frame with the optimal setpoints of the EV demand
* **sessions**: coordinated EV sessions' data frame according to the setpoint
* **log**: list with messages about the smart charging algorithm (flexibility needs and availability, sessions modified, etc.). This is empty by default and filled with messages when `include_log = TRUE`.



## Smart chargig for grid congestion

Imagine that we have to charge these EVs in an installation that has a **maximum grid connection of 8 kW**:

```{r}
grid_capacity <- 8
```

Since our peak goes above 18kW, we need to use smart charging to allow the EV users to charge under the grid capacity. In order to define a **grid capacity** in the `smart_charing()` function, a column with the same name than the EV user profile in the `sessions` parameter (in this example "HomeEV") must be found in the `opt_data` parameter. Therefore, a `"HomeEV"` column is added to the `opt_data` object as a setpoint for the EV user profile "HomeEV":

```{r}
opt_data <- tibble(
  datetime = ev_demand$datetime, # Same date-time sequence than the demand
  HomeEV = grid_capacity
)
head(opt_data)
```

Also, the parameter `opt_objective` must be set to `"none"` in order to skip optimization, since the setpoint for the smart charging algorithm will be given by the `opt_data` column named "HomeEV". 

Then, **which smart charging method should we use?** Let's compare them, considering optimization windows of 24 hours from 6:00AM to 6:00AM.


First we can try with the **postpone** strategy setting `method="postpone"`:

```{r}
sc_results <- ev_sessions %>% 
  smart_charging(
    opt_data = opt_data, 
    opt_objective = "none",
    method = "postpone",
    window_days = 1, 
    window_start_hour = 6
  )
```

```{r}
ev_demand_opt <- sc_results$sessions %>% 
  get_demand(by = "Session") 
```

```{r, echo = F}
ev_demand_opt %>% 
  dyplot(ylab = "Power demand (kW)", stackedGraph = T, fillGraph = T, legend_width = 150) %>% 
  dyLimit(grid_capacity, "Grid capacity", color = "red") %>% 
  dyAxis("y", valueRange = c(0, 10))
```
We see that the later EV sessions have been shifted during the night instead of start charging right at the connection time during peak hours. This is not a problem if all EVs are completely charged during their connection times, but it may suppose a risk for the late EV connections that have to wait until all previous sessions have been charged. This issue is solved with the **interrupt** strategy, where a rotation system is designed to charge a similar amount of energy to all connected EVs.

Let's see the difference with `method="interrupt"`:

```{r}
sc_results <- ev_sessions %>% 
  smart_charging(
    opt_data = opt_data, 
    opt_objective = "none",
    method = "interrupt",
    window_days = 1, 
    window_start_hour = 6
  )
```

```{r}
ev_demand_opt <- sc_results$sessions %>% 
  get_demand(by = "Session") 
```

```{r, echo = F}
ev_demand_opt %>% 
  dyplot(ylab = "Power demand (kW)", stackedGraph = T, fillGraph = T, legend_width = 150) %>% 
  dyLimit(grid_capacity, "Grid capacity", color = "red") %>% 
  dyAxis("y", valueRange = c(0, 10))
```

Now we see the same aggregated result but we also see that the individual sessions are interrupted to enable new sessions to charge, for a better equality among EV users. However, both postpone and interrupt methods do not provide high flexibility in terms of power since the individual charging power can't be reduced. In that sense, the **curtail** method provides a higher flexibility potential to adapt the EV demand to more variable or fluctuating signals such as flexibility markets or energy prices. 

Let's see the difference with `method="curtail"`:

```{r}
sc_results <- ev_sessions %>% 
  smart_charging(
    opt_data = opt_data, 
    opt_objective = "none",
    method = "curtail",
    window_days = 1, 
    window_start_hour = 6
  )
```

```{r}
ev_demand_opt <- sc_results$sessions %>% 
  get_demand(by = "Session") 
```

```{r, echo = F}
ev_demand_opt %>% 
  dyplot(ylab = "Power demand (kW)", stackedGraph = T, fillGraph = T, legend_width = 150) %>% 
  dyLimit(grid_capacity, "Grid capacity", color = "red") %>% 
  dyAxis("y", valueRange = c(0, 10))
```

Now we see that the EV demand is completely adapted to the grid capacity, sharing the full capacity among all different charging EVs in equal parts.


Imagine now that the **maximum grid connection has to be reduced to 4 kW**, so that we can charge only one EV at a time:

```{r}
grid_capacity <- 4
```

In this highly restricted scenario, applying any one of the previous smart charging methods is resulting in power profile that surpasses the grid capacity:

```{r, echo=F}
ev_demand %>% 
  aggregate_timeseries("original") %>% 
  mutate(
    postpone = ev_sessions %>% 
      smart_charging(
        opt_data = ev_demand %>% mutate(HomeEV = grid_capacity), 
        opt_objective = "none",
        method = "postpone",
        window_days = 1, 
        window_start_hour = 6
      ) %>% 
      .$sessions %>% get_demand() %>% .[[2]],
    interrupt = ev_sessions %>% 
      smart_charging(
        opt_data = ev_demand %>% mutate(HomeEV = grid_capacity), 
        opt_objective = "none",
        method = "interrupt",
        window_days = 1, 
        window_start_hour = 6
      ) %>% 
      .$sessions %>% get_demand() %>% .[[2]],
    curtail = ev_sessions %>% 
      smart_charging(
        opt_data = ev_demand %>% mutate(HomeEV = grid_capacity), 
        opt_objective = "none",
        method = "curtail",
        window_days = 1, 
        window_start_hour = 6
      ) %>% 
      .$sessions %>% get_demand() %>% .[[2]]
  ) %>% 
  dyplot(ylab = "Power demand (kW)", stackedGraph = F, fillGraph = T, legend_width = 150, strokeWidth = 2) %>% 
  dySeries("original", color = "gray", strokePattern = "dashed", fillGraph = F) %>% 
  dyLimit(grid_capacity, "Grid capacity", color = "red")
```
The demand is being postponed by the smart charging algorithm to match the setpoint during all possible hours, but when the end of the optimization window is reaching, the vehicle must charge to accomplish with the energy required by the EV user. This means that all EV users are not able to charge all their requirements under these grid conditions. 

There may be some scenarios where the charging point operator can assume this kind of impact to the user in order to mantain grid stability, so we can also consider that is not mandatory to fill the batteries of all vehicles. This can be configured with the parameter `energy_min`, as the required minimum ratio of the energy required by the EV user. Let's try a value of `energy_min = 0` to see how the algorithm works and the corresponding impact to the vehicles:

```{r, echo = FALSE}
energy_min <- 0
sc_results <- map(
  set_names(c("postpone", "interrupt", "curtail")),
  ~ ev_sessions %>% 
    smart_charging(
      opt_data = ev_demand %>% mutate(HomeEV = grid_capacity), 
      opt_objective = "none",
      method = .x,
      window_days = 1, 
      window_start_hour = 6,
      energy_min = energy_min
    )
)
```

```{r, echo=FALSE}
ev_demand %>% 
  aggregate_timeseries("original") %>% 
  mutate(
    postpone =  sc_results$postpone$sessions %>% 
      get_demand() %>% .[[2]],
    interrupt = sc_results$interrupt$sessions %>% 
      get_demand() %>% .[[2]],
    curtail =  sc_results$curtail$sessions %>% 
      get_demand() %>% .[[2]],
  ) %>% 
  dyplot(ylab = "Power demand (kW)", stackedGraph = F, fillGraph = T, legend_width = 150, strokeWidth = 2) %>% 
  dySeries("original", color = "gray", strokePattern = "dashed", fillGraph = F) %>% 
  dyLimit(grid_capacity, "Grid capacity", color = "red")
```


Now we see that the grid capacity constraint is respected, **but at which price?**

Let's see the impact on the EV user in terms of **percentage of energy charged**:

```{r, echo=FALSE}
sc_results$postpone$sessions %>%
  group_by(Session) %>% summarise(Energy = sum(Energy)) %>% 
  left_join(ev_sessions %>% select(Session, EnergyRequired = Energy), by = "Session") %>% 
  mutate(EnergyCharged = round(Energy/EnergyRequired*100)) %>% 
  select(Session, Postpone = EnergyCharged) %>% 
  left_join(
    sc_results$interrupt$sessions %>%
      group_by(Session) %>% summarise(Energy = sum(Energy)) %>% 
      left_join(ev_sessions %>% select(Session, EnergyRequired = Energy), by = "Session") %>% 
      mutate(EnergyCharged = round(Energy/EnergyRequired*100)) %>% 
      select(Session, Interrupt = EnergyCharged),
    by = "Session"
  ) %>% 
  left_join(
    sc_results$curtail$sessions %>%
      group_by(Session) %>% summarise(Energy = sum(Energy)) %>% 
      left_join(ev_sessions %>% select(Session, EnergyRequired = Energy), by = "Session") %>% 
      mutate(EnergyCharged = round(Energy/EnergyRequired*100)) %>% 
      select(Session, Curtail = EnergyCharged),
    by = "Session"
  ) %>% 
  bind_rows(
    tibble(Session = "Mean", Postpone = round(mean(.$Postpone)),
           Interrupt = round(mean(.$Interrupt)), Curtail = round(mean(.$Curtail)))
  ) %>% 
  knitr::kable(caption = "Percentage of energy charged (%)")
```

We see that the curtail strategy is the one with lowest average impact on the users, in contrast to postpone which has the highest impact, specially in late charging sessions.


## Smart charging for optimization

Another possible application of smart charging is to coordinate EV sessions to maximize the use of renewable energy or to minimize the energy cost. Let's see some examples of these optimizations using the example data set `energy_profiles` provided by `flextools`.

```{r}
head(energy_profiles)
```

To this data set with building demand, solar PV production and energy prices, some EV demand can be added using the `evsim` package.

First, we will filter the example `energy_profiles` data set for just a single week, rename the variables with custom names to fit the column names required in the `flextools` functions, and increase the magnitude of the energy variables to coincide with the order of the building demand:

```{r}
opt_data <- energy_profiles %>% 
  filter(isoweek(datetime) == 18, wday(datetime, week_start = 1) <= 5) %>% 
  rename(
    production = solar
  ) %>% 
  mutate(
    production = production*4,
    building = building*2
  )
```


Now we can simulate EV sessions corresponding to our custom EV model during the first 3 days:

```{r}
set.seed(1234)
ev_sessions <- simulate_sessions(
  evmodel = ev_model, 
  sessions_day = tibble(time_cycle = c("Workdays", "Weekends"), n_sessions = c(5, 5)),
  user_profiles = NULL,
  charging_powers = tibble(power = 3.7, ratio = 1), 
  dates = unique(date(opt_data$datetime))[1:3], 
  resolution = 15
) %>% 
  filter(Profile == "HomeEV")
```

We can calculate the aggregated EV demand with function `evsim::get_demand()`:

```{r}
ev_demand <- get_demand(ev_sessions)
```

```{r, echo = FALSE}
ev_demand %>% 
  dyplot(ylab = "Power demand (kW)", fillGraph = T, strokeWidth = 2)
```

And add the EV demand on top of the other energy profiles:

```{r, echo=FALSE}
opt_data %>% 
  left_join(ev_demand, by = "datetime") %>% 
  select(-starts_with("price")) %>% 
  dyplot() %>% 
  dyAxis("y", "Power (kW)") %>% 
  dySeries("production", "Production", "orange") %>% 
  # dySeries("building", "Building", "navy") %>%
  # dySeries("HomeEV", "Electric vehicles", "brown") %>% 
  dyStackedRibbonGroup(c("HomeEV","building"), color = c("brown", "navy")) %>% 
  dyOptions(strokeWidth = 2)
```

We see that the EV demand cause an even higher demand peak during the evening peak hours. To solve this issue, the EV sessions can be coordinated with the buildig demand profile to obtain a lower aggregated peak. This can be done with function `smart_charging()`, using a net power optimization (i.e. `opt_objective = "grid"`, see more information in [Net power optimization article](https://mcanigueral.github.io/flextools/articles/minimize_net_power.html), a `"curtail"` method, considering optimization windows of 24 hours from 6:00AM to 6:00AM, and renaming the variable `"building"` to `"static"` in the `opt_data` object:

```{r}
sc_results <- ev_sessions %>% 
  smart_charging(
    opt_data = opt_data %>% rename(static = building), 
    opt_objective = "grid",
    method = "curtail",
    window_days = 1, 
    window_start_hour = 6
  )
```

If we check the **setpoints** object from the `smart_charging()` results, is a data frame with the optimal EV demand:

```{r}
head(sc_results$setpoints)
```

```{r, echo = FALSE}
ev_demand %>% 
  rename(original = HomeEV) %>% 
  left_join(
    sc_results$setpoints %>% rename(setpoint = HomeEV), by = "datetime"
  ) %>% 
  dyplot(ylab = "Power demand (kW)", fillGraph = T, strokeWidth = 2) %>% 
  dySeries("original", "HomeEV Original", color = "gray", strokePattern = "dashed") %>% 
  dySeries("setpoint", "HomeEV setpoint", color = "black", strokePattern = "dashed")
```

The optimal setpoint is a specific power curve to obtain the flattest possible net power curve. We can check that by adding these EV profiles to the building (`consumption = building + setpoint`) and calculating the net power profile with function `get_energy_balance()`:

```{r, echo = F}
opt_data %>% 
  select(datetime, building, production) %>% 
  left_join(
    sc_results$setpoints %>% rename(setpoint = HomeEV), by = "datetime"
  ) %>% 
  mutate(
    consumption = building + setpoint
  ) %>% 
  get_energy_balance() %>% 
  select(datetime, consumption, production, net) %>% 
  dyplot(ylab = "Power demand (kW)", fillGraph = T, strokeWidth = 2) %>% 
  dySeries("production", "Production", color = "orange") %>% 
  dySeries("consumption", "Consumption",  color = "navy") %>%
  dySeries("net", "Net power", color = "brown")
```

From this graph we see that during the evening peak, the demand can be completely flat, shaving the peak, if the EVs can be coordinated to match the setpoint. Whether the EVs can provide this flexibility or not is shown by the coordinated sessions data set provied by `smart_charging()` function:

```{r}
ev_demand_opt <- sc_results$sessions %>% 
  get_demand() 
```

```{r, echo = F}
ev_demand %>% 
  rename(original = HomeEV) %>% 
  left_join(
    sc_results$setpoints %>% rename(setpoint = HomeEV), by = "datetime"
  ) %>% 
  left_join(
    ev_demand_opt %>% rename(scheduled = HomeEV), by = "datetime"
  ) %>% 
  dyplot(ylab = "Power demand (kW)", legend_width = 200, strokeWidth = 2) %>% 
  dySeries("original", "Original HomeEV", color = "gray", strokePattern = "dashed") %>% 
  dySeries("setpoint", "Optimal setpoint", color = "black", strokePattern = "dashed") %>% 
  dySeries("scheduled", "Scheduled HomeEV", color = "navy", fillGraph = T)
```

We see that during some timeslots the EV demand surpasses the setpoint. This happens in order to charge all energy requirements for all EVs, since by default `energy_min = 1`. If we set `energy_min = 0.5`, we see that we can meet the setpoint to all time slots but not all energy required is charged to the vehicles.

```{r}
sc_results <- ev_sessions %>% 
  smart_charging(
    opt_data = opt_data %>% rename(static = building), 
    opt_objective = "grid",
    method = "curtail",
    window_days = 1, 
    window_start_hour = 6,
    energy_min = 0.5
  )
```

```{r, echo = F}
ev_demand_opt <- sc_results$sessions %>% 
  get_demand()
ev_demand %>% 
  rename(original = HomeEV) %>% 
  left_join(
    sc_results$setpoints %>% rename(setpoint = HomeEV), by = "datetime"
  ) %>% 
  left_join(
    ev_demand_opt %>% rename(scheduled = HomeEV), by = "datetime"
  ) %>% 
  dyplot(ylab = "Power demand (kW)", legend_width = 200, strokeWidth = 2) %>% 
  dySeries("original", "Original HomeEV", color = "gray", strokePattern = "dashed") %>% 
  dySeries("setpoint", "Optimal setpoint", color = "black", strokePattern = "dashed") %>% 
  dySeries("scheduled", "Scheduled HomeEV", color = "navy", fillGraph = T)
```

Another solution to avoid surpassing the setpoint but charging all energy required by EVs is to configure a certain power threshold between the EV demand and the optimization setpoint (`power_th` parameter in `smart_charging()`). With this parameter, we assume that the EV demand is acceptable between a margin above and below the setpoint. Let's see the result with a power threshold of a 10%:


```{r}
sc_results <- ev_sessions %>% 
  smart_charging(
    opt_data = opt_data %>% rename(static = building), 
    opt_objective = "grid",
    method = "curtail",
    window_days = 1, 
    window_start_hour = 6,
    power_th = 0.1
  )
```

```{r, echo = F}
ev_demand_opt <- sc_results$sessions %>% 
  get_demand()
ev_demand %>% 
  rename(original = HomeEV) %>% 
  left_join(
    sc_results$setpoints %>% rename(setpoint = HomeEV), by = "datetime"
  ) %>% 
  left_join(
    ev_demand_opt %>% rename(scheduled = HomeEV), by = "datetime"
  ) %>% 
  dyplot(ylab = "Power demand (kW)", legend_width = 200, strokeWidth = 2) %>% 
  dySeries("original", "Original HomeEV", color = "gray", strokePattern = "dashed") %>% 
  dySeries("setpoint", "Optimal setpoint", color = "black", strokePattern = "dashed") %>% 
  dySeries("scheduled", "Scheduled HomeEV", color = "navy", fillGraph = T)
```







Now we can repeat the same with the other smart charging strategies:

```{r, echo = FALSE}
sc_results <- map(
  set_names(c("postpone", "interrupt", "curtail")),
  ~ ev_sessions %>% 
    smart_charging(
    opt_data = opt_data %>% rename(static = building), 
      opt_objective = "grid",
      method = .x,
      window_days = 1, 
      window_start_hour = 6,
      power_th = 0.1
    )
)
```

```{r, echo=FALSE}
ev_demand %>% 
  aggregate_timeseries("original") %>% 
  left_join(
    sc_results$postpone$setpoints %>% rename(setpoint = HomeEV), 
    by = "datetime"
  ) %>% 
  mutate(
    postpone =  sc_results$postpone$sessions %>% 
      get_demand() %>% .[[2]],
    interrupt = sc_results$interrupt$sessions %>% 
      get_demand() %>% .[[2]],
    curtail =  sc_results$curtail$sessions %>% 
      get_demand() %>% .[[2]],
  ) %>% 
  dyplot(ylab = "Power demand (kW)", legend_width = 200, strokeWidth = 2) %>% 
  dySeries("original", "Original HomeEV", color = "gray", strokePattern = "dashed") %>% 
  dySeries("setpoint", "Optimal setpoint", color = "black", strokePattern = "dashed")
```


Making zoom at a single day, for example the evening of 2nd May, we can see that the *postpone* strategy can't fully shift the power peak from peak hours, since all sessions that have already started charging are not flexible. On the other hand, *interrupt* and *curtail* strategies are better in peak shaving since more EV sessions are candidates to provide flexibility. 

At the same time, the *interrupt* method limits EV charging below setpoint since it can only stop sessions. This results in a lack of flexibility at the end of the optimization window and a rebound effect since the sessions must charge before the end of connection time.

<!-- Finally, even though all sessions have been coordinated in a different way, the total energy consumed by all EV users must remain the same: -->

<!-- ```{r} -->
<!-- sum(ev_sessions$Energy) -->
<!-- sum(sc_results$postpone$sessions$Energy) -->
<!-- sum(sc_results$interrupt$sessions$Energy) -->
<!-- sum(sc_results$curtail$sessions$Energy) -->
<!-- ``` -->


